---
title: "DDSAnalytics"
author: "Travis R. Daun, Joe Jiang, Karl J. Jurek"
date: "February 21, 2019"
output: html_document
---

```{r stuff, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=10,fig.height=11)
```

```{r setup, include=T, echo=TRUE}
library(httr)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)

# Reading in the 'CaseStudy2-data.csv' data file
attrition=read.csv("https://raw.githubusercontent.com/dauntr/MSDS6306_JKT_Case_Study_2/master/datasets/CaseStudy2-data.csv", header=T, stringsAsFactors=F)

# Validating there is no missing data
sapply(attrition,function(x) sum(is.na(x)))

# Changing various columns to 'factors'
attrition$BusinessTravel=as.factor(attrition$BusinessTravel)
attrition$Attrition=as.factor(attrition$Attrition)
attrition$Department=as.factor(attrition$Department)
attrition$EducationField=as.factor(attrition$EducationField)
attrition$Gender=as.factor(attrition$Gender)
attrition$JobRole=as.factor(attrition$JobRole)
attrition$MaritalStatus<- as.factor(attrition$MaritalStatus)
attrition$Over18=as.factor(attrition$Over18)
attrition$OverTime=as.factor(attrition$OverTime)
attrition$Education=as.factor(attrition$Education)
attrition$EnvironmentSatisfaction=as.factor(attrition$EnvironmentSatisfaction)
attrition$JobInvolvement=as.factor(attrition$JobInvolvement)
attrition$JobLevel=as.factor(attrition$JobLevel)
attrition$JobSatisfaction=as.factor(attrition$JobSatisfaction)
attrition$PerformanceRating=as.factor(attrition$PerformanceRating)
attrition$RelationshipSatisfaction=as.factor(attrition$RelationshipSatisfaction)
attrition$StockOptionLevel=as.factor(attrition$StockOptionLevel)
attrition$WorkLifeBalance=as.factor(attrition$WorkLifeBalance)

# Remove the 'over18' column because it only has one level (all 'yes')
attrition$Over18=NULL  

# Fit full logistic model 
fullmodel=glm(Attrition~., family= binomial(link="logit"), data=attrition)

# Looking at the 'fullmodel'
summary(fullmodel)
```

|
|
# Introduction
|
|
|         DDS Analytics has been charged with the task of looking into the top reasons why personnel leave the company.  
|         We have performed various analysis with the data set provided to come to the conclusions requested.

|
|
## Attrition Rates by Department
|
|         The following graphs display the attrition rates by department.

|
|
|
```{r Leave_Dept, fig.width=10,fig.height=7, echo=TRUE}

# Graphs for attrition by Dept
attrition%>%ggplot() +
  geom_bar(aes(Department, fill= Attrition), color= 'black', position="stack")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='Department Attrition by Count')+
  theme(plot.title=element_text(hjust=0.5))+
  ylab('Count')+
  theme(axis.text.x=element_text(angle=60, hjust=1))

attrition%>%ggplot(aes(Department, fill= Attrition)) +
  geom_bar(aes(y=(..count..)/sum(..count..)), color ='black', position="fill")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='Department Attrition Rate by Percentage')+
  theme(plot.title=element_text(hjust=0.5))+
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  ylab('Percentages')+
  geom_hline(yintercept=0.1612, color="blue")
```

|
|
## Attrition Rates by Job Role
|
|         The following graphs display the attrition rates by job role.

|
|
|
```{r Leave_Job_Role, fig.width=10,fig.height=7, echo=TRUE}

# Graphs for attrition by Job Roles
attrition%>%ggplot() +
  geom_bar(aes(JobRole, fill= Attrition), color='black', position="stack")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='The Effect of Job Role on Attrition by Count')+
  theme(plot.title=element_text(hjust=0.5))+
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  ylab('Count')+
  xlab('Job Role')

attrition%>%ggplot(aes(JobRole, fill= Attrition)) +
  geom_bar(aes(y=(..count..)/sum(..count..)), color='black',position="fill")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='The Effect of Job Role on Attrition Rate by Percentage')+
  theme(plot.title=element_text(hjust=0.5))+
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  ylab('Percentages')+
  xlab('Job Role')+
  geom_hline(yintercept=0.1612, color="blue")
```

|
|
## Top Three Reasons Employees Leave
|
|         1. Overtime

|
|
|
```{r Leave_Overtime, fig.width=10,fig.height=7, echo=TRUE}
# Graphs for attrition by Overtime
attrition%>%ggplot() +
  geom_bar(aes(OverTime, fill= Attrition), color='black', position="stack")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  theme(plot.title=element_text(hjust=0.5))+
  ylab('Count')+
  xlab('Overtime')+
  ggtitle(label='The Effect of Overtime on Attrition by Count')

attrition%>%ggplot(aes(OverTime, fill= Attrition)) +
  geom_bar(aes(y=(..count..)/sum(..count..)), color='black',position="fill")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='The Effect of Overtime on Attrition Rate by Percentage')+
  theme(plot.title=element_text(hjust=0.5))+
  ylab('Percentages')+
  xlab('Overtime')+
  geom_hline(yintercept=0.1612, color="blue")
```

|
|         2. Job Involvement

|
|
| 
```{r Leave_Job_Involvement, fig.width=10,fig.height=7, echo=TRUE}

# Graphs for attrition by Job Involvement
attrition%>%ggplot() +
  geom_bar(aes(x=JobRole, fill=JobInvolvement ), color='black', position="stack")+
  scale_fill_manual(values=c("1"="red","2"="dark green", '3'='blue', '4'='yellow'))+
  theme(plot.title=element_text(hjust=0.5))+
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  ggtitle(label='Employee Involvement Scores Across All Roles by Count')+
  ylab('Count')+
  xlab('Job Role')

attrition%>%ggplot(aes(JobRole, fill=JobInvolvement )) +
  geom_bar(aes(y=(..count..)/sum(..count..)),color ='black', position="fill")+
  scale_fill_manual(values=c("1"="red","2"="dark green", '3'='blue', '4'='yellow'))+
  ggtitle(label='Employee Involvement Scores Across All Roles by Percentage')+
  theme(plot.title=element_text(hjust=0.5))+
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  ylab('Percentages')+
  xlab('Job Role')
```

|
|         3. Frequent Travel

|
|
| 
```{r Leave_Job_Travel, fig.width=10,fig.height=7, echo=TRUE}

# Graphs for attrition by Frequent Travel
attrition%>%ggplot() +
  geom_bar(aes(BusinessTravel, fill= Attrition), color='black', position="stack")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='The Effect of Business Travel on Attrition by Count')+
  theme(plot.title=element_text(hjust=0.5))+
  ylab('Count')+
  xlab('Travel')

attrition%>%ggplot(aes(BusinessTravel, fill= Attrition)) +
  geom_bar(aes(y=(..count..)/sum(..count..)), color='black', position="fill")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='The Effect of Business Travel on Attrition Rate by Percentage')+
  theme(plot.title=element_text(hjust=0.5))+
  ylab('Percentages')+
  geom_hline(yintercept=0.1612, color="blue")+
  ylab('Percentages')+
  xlab('Travel')
```

|
|
## Top Reasons Why Employees Stay
|
|         1. Stock Options

|
|
|
```{r Stay_Stock_Options, fig.width=10,fig.height=7, echo=TRUE}

# Graphs for Staying because of Stock Options
attrition%>%ggplot() +
  geom_bar(aes(StockOptionLevel, fill= Attrition), position="stack")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='The Effect of Stock Option Levels on Attrition by Count')+  
  theme(plot.title=element_text(hjust=0.5))+
  ylab('Count')+
  xlab('Stock Options Level')


attrition%>%ggplot(aes(StockOptionLevel, fill= Attrition)) +
  geom_bar(aes(y=(..count..)/sum(..count..)), position="fill")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='The Effect of Stock Option Levels on Attrition Rate by Percentage')+
  theme(plot.title=element_text(hjust=0.5))+
  ylab('Percentages')+
  geom_hline(yintercept=0.1612, color="blue")+
  ylab('Percentage')+
  xlab('Stock Options Level')
```

|
|         2. At the Company for at least three years.

|
|
| 
```{r Stay_Company_Time, fig.width=10,fig.height=7, echo=TRUE}

# Graphs for Staying because of being with the company greater than three years
attrition%>%ggplot() +
  geom_histogram(aes(x=YearsAtCompany, fill= Attrition), bins =41, color='black', position="stack")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='The Effect of Job Involvement on Attrition by Count')+
  theme(plot.title=element_text(hjust=0.5))+
  ylab('Count')+
  xlab('Years with Company')

attrition%>%ggplot(aes(x=YearsAtCompany, fill= Attrition)) +
  geom_histogram(aes(y=(..count..)/sum(..count..)), bins =41, color ='black', position="fill")+
  scale_fill_manual(values=c("Yes"="red","No"="dark green"))+
  ggtitle(label='The Effect of Job Involvement on Attrition Rate by Percentage')+
  theme(plot.title=element_text(hjust=0.5))+
  ylab('Percentages')+
  xlab('Years with Company')+
  geom_hline(yintercept=0.1612, color="blue")
```

|
|
## Conclusion and Recommendations
|
|         1. Overtime is the most influential factor in why an employee leaves the company.  You should find ways to 
|            improve work/life balance for your employees.
|         2. If the employee does not feel a sense of involvment they will leave.  A strategy should be developed to keep |            employees engaged.
|         3. Reduce travel time for employees as much as possible. This has a two-fold effect.  First, you wil retain
|            more of your employees.  Second, you will reduce expenses for the company.
|         4. The largest group of employees are the young and new employees.  If you are able to keep to retain the
|            employee past three years they will be with the company for a long period.  Ways of accomplishing this are:
|                     A. Mentoring and more time with their manager.
|                     B. Offering stock options earlier or immediately when hired.

|
|
## Appendix

|
|         1. Naive Bayes Model

|
|
| 
```{r Appendix_Naive_Bayes, fig.width=10,fig.height=7, echo=TRUE}

# Installing package and library 'e1071' for Naive_Bayes
install.packages("e1071", dependencies=F, repos="http://cran.us.r-project.org")
library(e1071)

# Creating the Naive-Bayes model
Naive_Bayes_Model=naiveBayes(Attrition ~., data=attrition)

# Viewing the Naive_Bayes model
Naive_Bayes_Model

# Creating the Naive-Bayes model predictions
NB_Predictions=predict(Naive_Bayes_Model,attrition)

# Viewing the Naive_Bayes model predictions
# NOTE: Confusion matrix to check accuracy
table(NB_Predictions, attrition$Attrition)
```

|
|         2. Trees

|
|
| 
```{r Appendix_Tree, fig.width=10,fig.height=7, echo=TRUE}

# Installing package and library 'rpart' and 'rpart.plot' for tree view
install.packages("rpart", dependencies=F, repos="http://cran.us.r-project.org")
install.packages("rpart.plot", dependencies=F, repos="http://cran.us.r-project.org")
library(rpart)

# Creating the tree
tree <- rpart(Attrition ~ Age + Education + MonthlyIncome + JobLevel + StockOptionLevel + MaritalStatus +
                 Department + BusinessTravel + DistanceFromHome + EducationField + Gender + JobRole +
                 NumCompaniesWorked + OverTime + PercentSalaryHike + TotalWorkingYears + TrainingTimesLastYear +
                 YearsAtCompany + YearsInCurrentRole + YearsSinceLastPromotion + YearsWithCurrManager + 
                 EnvironmentSatisfaction + JobInvolvement + JobSatisfaction + PerformanceRating +
                 RelationshipSatisfaction + WorkLifeBalance,
                 data=attrition,  method="class")

# Viewing the summary of 'tree'
summary(tree)

#Calling the library 'rpart.plot'for plotting 'tree'
library(rpart.plot)

# Plotting 'tree'
rpart.plot(tree)
```